Clustering analysis and statistics

### 2019-05-05 notebooks/Clustering analysis.Rmd

# Allele distance distribution
```{r}
# calculates distances from cgMLST data and returns
# the lower triangular of the matrix, with (diag = FALSE)
# or without (diag = TRUE) the diagonal
lower_tri_dist_calc <- function(data, diag = TRUE, data_frame = FALSE) {
  dist <- as.matrix(daisy(data, metric = "gower"))
  if (diag == TRUE) {
    dist[upper.tri(dist, diag = TRUE)] <- NA
  } else {
    dist[upper.tri(dist, diag = FALSE)] <- NA
  }
  dist_vec <- as.vector(as.matrix(dist))
  dist_compl <- dist_vec[!is.na(dist_vec)]

@@

# Distribution of iterated distances
Distribution of percentages of distances < 0.1 or > 0.8. The distribution is based on 1000 iterations of the distance data. Observed values are denoted as arrows for each animal species; dark blue = Broiler, light blue = Pig, red = Red fox, yellow = Wild bird. The mean value is denoted as a vertical black line.
```{r, warning=FALSE, message=FALSE, dpi=300}
# generate data frame from allele data
alleles <- cgMLST_allele_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         new_id = case_when(species == "Broiler" ~ paste("BR", n, sep = "_"),
                            species == "Pig" ~ paste("PI", n, sep = "_"),
                            species == "Red fox" ~ paste("RF", n, sep = "_"),
                            species == "Wild bird" ~ paste("WB", n, sep = "_"))) %>%
  ungroup() %>%
  select(new_id, everything(), -c(species, id)) %>%
  mutate_all(funs(as.factor)) %>%
  column_to_rownames("new_id")

@@

# functions used in iteration analyses

@@

# Run functions
summary_stats <- run_df %>%
  mutate(total_mean = round(mean(value), 3)) %>%
  group_by(group1) %>%
  mutate(mean = round(mean(value), 3),
         var = round(var(value), 3)) %>%
  select(group1, total_mean, mean, var) %>%
  summarise_all(funs(func_paste))

@@

summary_stats

>>>>

# Allele distance distribution
Basic statistics are presented in the table.
```{r, message=FALSE, warning=FALSE}
# calculates distances from cgMLST data and returns
# the lower triangular of the matrix, with (diag = FALSE)
# or without (diag = TRUE) the diagonal
lower_tri_dist_calc <- function(data, diag = TRUE, data_frame = FALSE) {
  dist <- as.matrix(daisy(data, metric = "gower"))
  if (diag == TRUE) {
    dist[upper.tri(dist, diag = TRUE)] <- NA
  } else {
    dist[upper.tri(dist, diag = FALSE)] <- NA
  }
  dist_vec <- as.vector(as.matrix(dist))
  dist_compl <- dist_vec[!is.na(dist_vec)]

@@

# generate data frame from allele data
alleles <- cgMLST_allele_data %>%
  left_join(isolate_data[,c("id","species")], by = "id") %>%
  group_by(species) %>%
  mutate(n = 1:n(),
         new_id = case_when(species == "Broiler" ~ paste("BR", n, sep = "_"),
                            species == "Pig" ~ paste("PI", n, sep = "_"),
                            species == "Red fox" ~ paste("RF", n, sep = "_"),
                            species == "Wild bird" ~ paste("WB", n, sep = "_"))) %>%
  ungroup() %>%
  select(new_id, everything(), -c(species, id)) %>%
  mutate_all(funs(as.factor)) %>%
  column_to_rownames("new_id")

@@

run_df %>%
  mutate(total_mean = round(mean(value), 3)) %>%
  group_by(group1) %>%
  mutate(mean = round(mean(value), 3),
         var = round(var(value), 3)) %>%
  select(group1, total_mean, mean, var) %>%
  summarise_all(funs(func_paste))
```

# Distribution of iterated distances
Distribution of percentages of distances < 0.1 or > 0.8. The distribution is based on 1000 iterations of the distance data. Observed values are denoted as arrows for each animal species; dark blue = Broiler, light blue = Pig, red = Red fox, yellow = Wild bird. The mean value is denoted as a vertical black line.
```{r, warning=FALSE, message=FALSE, dpi=300}
# functions used in iteration analyses

@@

# Run functions

@@
